---
title: "CN334 Data Analytics"
author: "Aakriti Sharma"
date: "2026-03-23"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 4
# function to calculate the means of a vector
```{r}
myMean <- function(y1) {
  mean <- sum(y1)/length(y1)  
  return(mean)                
}
# Test the function
testVec <- rnorm(50, 20, 4)  
mean(testVec) 
myMean(testVec)               
```
# calculate t-test statistic
```{r}
twosam <- function(y1, y2) {
  n1 <- length(y1)          
  n2 <- length(y2)          
  yb1 <- mean(y1)           
  yb2 <- mean(y2)            
  s1 <- var(y1)             
  s2 <- var(y2)             
  
  s <- ((n1 - 1) * s1 + (n2 - 1) * s2)/(n1 + n2 - 2)  
  tst <- (yb1 - yb2)/sqrt(s * (1/n1 + 1/n2))         
  
  return(tst)               
}
```
# conditional statement
# if statement
```{r}
x <- 9

if (x > 0) sqrt(x) else sqrt(-x)
# Vectorized version
ifelse(x >= 0, sqrt(x), sqrt(-x))
```
# switch
```{r}
spread <- function(x, type) {
  switch(type,
         sd = sd(x),              
         mad = mad(x),           
         IQR = IQR(x)/1.349)     
}
samp <- rnorm(50)
spread(samp, 2)
spread(samp,"IQR")
```
# Repetitive execution
# Forloops and while loops
```{r}
# For loop
for (i in 1:5) print(i^2)
# While loop 
x <- 1
y <- 16

while (x^2 < y) {
  cat(x, "squared is", x^2, "\n")   
  x <- x + 1                        
}
```

# Viewing Function Code
```{r}
# View arguments
args(IQR)
# View function body
body(IQR)
xx <- sample(1:30,10)
quantile(xx)
IQR(xx)
```
# methods to view the function
```{r}
#visible
mean
methods(mean)
mean.default
# Non visible
`?`(t.test)
t.test
methods(t.test)
```
# To reduce the output
```{r}
head((stats:::t.test.default), 5)
print("truncated...")
tail((stats:::t.test.default),10)
```
# Apply function
```{r}
# Apply function on dataset
summary(women)
# Column means
colMeans(women)
# Apply mean column-wise
apply(women, 2, mean)
# Check equality
testEq <- all(rowMeans(women) == apply(women, 1, mean))
print(testEq)
if(testEq) print("rowMeans is equivalent to apply(df,1,mean)")
```
# create a function using apply
```{r}
colSd <- function(df) apply(df,2,sd)
colSd(women)
```

# tapply
```{r}
# Create groups
ageSplit <- ifelse(women$weight < 35, "under35", "over35")
print(ageSplit)
# Apply function by group
tapply(women$weight, ageSplit, length)
tapply(women$weight, ageSplit, summary)
```

# lapply and sapply
```{r}
# Load built-in dataset
data(ToothGrowth)   
# Create list
myList <- list(
  ToothGrowth = ToothGrowth,      
  WomenAge = women$age,         
  beta = exp(-3:3),               
  logicalVec = c(TRUE, FALSE, FALSE, TRUE)  
)

# Apply length function using lapply
res1 <- lapply(myList, length)
print(res1)

# Print class of result
print(paste("Class of res1:", class(res1)))

# Simplified version using sapply
res2 <- sapply(myList, length)
print(res2)

# Print class of result
print(paste("Class of res2:", class(res2)))
```

# Efficient for loop in R(use apply)
```{r}
myList <- list(WomenMat = women, WomenAge = women$age, beta = exp(-3:3),
               logicalVec = c(TRUE, FALSE, FALSE, TRUE))

# compute the list mean for each element
res1 <- lapply(myList, length)

print(res1)
print(paste("Class of res1:", class(res1)))

res2 <- sapply(myList, length)

print(res2)
print(paste("Class of res2:", class(res2)))
```

# use built-in function
```{r}
# Create large matrix
myMA <- matrix(rnorm(1e+06), 1e+05, 10, dimnames = list(1:1e+05, paste("V", 1:10, sep="")))

results <- NULL
system.time(for (i in seq(along = myMA[,1])) results <- c(results, mean(myMA[i, ])))

results <- numeric(length(myMA[,1]))
system.time(for (i in seq(along = myMA[,1])) results[i] <- mean(myMA[i, ]))

system.time(myMAmean <- apply(myMA, 1, mean))
system.time(myMAsd <- apply(myMA, 1, sd))
system.time(myMAsd <- sqrt((rowSums((myMA - rowMeans(myMA))^2))/
                           (length(myMA[,1]) - 1)))
```
# Function for parsing text
```{r}
# grep
grep("A", LETTERS)
grep("A", LETTERS, value = TRUE)
#sub()
# Substitute characters within a string
sub("B", "A", LETTERS)
# strsplit
# Split a character string
a <- date()
strsplit(a, " ")
strsplit(a, "J ")
b <- strsplit(a, " 11")
class(b)
b <- unlist(b)
class(b)
# For special characters use double backslash
a <- "aedin@jimmy.harvard.edu"
strsplit(a, "\\.")
```
# missing values
```{r}
# Create vector with NA values
a <- c(1:10, NA, NA)
a <- c(1:10, NA, NA)
# Summary shows statistics and number of NA values
summary(a)
# Mean without handling NA
mean(a)
# Mean after removing NA values
mean(a, na.rm = TRUE)
```
# writing Function
```{r}
# default values
prob <- c(0.5, 0.9, 0.95, 0.975, 0.99)
args (qnorm)
qnorm (prob)
qnorm(prob, 2)
qnorm(prob, mean = 2, sd = 1)

#positional arguments
prob <- c(0.5, 0.9, 0.95, 0.975, 0.99)
args (qnorm)
qnorm (p = prob, mean = 0, sd = 1, lower.tail = TRUE, log.p =
FALSE)
qnorm(prob, 0, 1, log = FALSE, low = TRUE)
qnorm(prob, 0, 1, TRUE, FALSE)
```
# scoping
```{r}
fn <- function(x) {
y <- 2 * x
print (x)
print (y)
print (z)
}

z <- 2
x <- 4
fn(x = 2)
```
# Lexical scope
```{r}
cube <- function (n) {
sq <- function () n * n
n * sq()
}
cube (2)
n <- 4
cube (2)
```

# System time
```{r}
df <- matrix(rnorm(5e+06), ncol = 20000)
system.time (apply (df, 1, mean))
system.time (rowMeans (df) )
```