---
title: "CN334 Data Analytics lab"
author: "Aakriti Sharma"
date: "2026-03-23"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 2
# Vectors
```{r cars}
# Numeric
vec1 <- 1:10
vec1
# Character
vec2 <- LETTERS[1:10]
vec2
# Logical
vec3 <- vec2 == "D"
vec3
```
# Matrix
```{r}
mat1 <- matrix(vec1, ncol = 2, nrow = 5)
print(mat1)
dim(mat1)
colnames(mat1) <- c("A", "B")
rownames(mat1) <- paste("N", 1:5, sep="")
print(mat1)
```

# List
```{r}
a <- 20
newList1 <- list(a, vec1, mat1)
print(newList1)
newList1 <- list(a = a, myVec = vec1, mat = mat1)
print(newList1)
```

# Data Frame
```{r}
df1 <- as.data.frame(mat1)
df1
```


# Factor
```{r}
charVec <- rep(LETTERS[1:3], 10)
table(charVec)
fac1 <- factor(charVec)
print(fac1)
attributes(fac1)
levels(fac1)
```

# array
```{r}
array(1:24, dim = c(2, 4, 3))

array(1:23, dim = c(2,4,3))

array(1:23, dim = c(2,4,3), dimnames = list(paste("Patient",1:2, sep = ""), LETTERS[1:4],c("X","Y","Z")))

```

# Attributes of R Objects
# Basic attributes
```{r}
# "numeric"
x <- 3
mode(x) 
# "character"
x <- "apple"
mode(x)
# integer
x <- 3.145
x + 2
# False, Logical
x==2
x <- x==2
x
mode(x)
```

# All R objects have mode

```{r}
# vectors of different modes Numeric
x <- 1:10
mode(x)
x <- matrix(rnorm(50), nrow=5, ncol=10)
mode(x)
# Character
x <- LETTERS[1:5]
mode(x) 
# Logical
x <- x == "D"
mode(x)
```

# Quick Exercise(Reapeat above,find the length and class in each case.)
```{r}
# Numeric vector
x <- 1:10          
length(x)         
class(x)  
x <- matrix(rnorm(50), nrow = 5, ncol =10)
length(x)
class(x)
# Character vector
x <- LETTERS[1:5] 
length(x)          
class(x)          
# Logical vector
x <- x == "D"      
length(x)          
class(x)         
```

# Dimension
```{r}
x <- matrix(5:14, nrow=2, ncol=5)
x
attributes(x)
```

# Creating and accessing objects
# Create vectors, matrices and data frames using seq,rep,rbind and cbind.
```{r}
# Create Vector
x.vec <- seq(1, 7, by = 2)
names(x.vec) <- letters[1:4]
# Create Matrices
xMat <- cbind(x.vec, rnorm(4), rep(5,4))
yMat <- rbind(1:3, rep(1,3))
z.mat <- rbind(xMat, yMat)
# Data Frame
x.df <- as.data.frame(xMat)
names(x.df) <- c("ind", "random", "score")
```

# Accessing Elemnts
```{r}
# Access first element of 'x.vec'
x.vec[1]
# or if you know the name 
x.vec["a"]
# Access an element of 'xMat' in the second row, third column
xMat[2,3]
# Display the seconds and third columns of matrix 'xMat'
xMat[,c(2:3)]
#or
xMat[,-c(1)]
```

# Access using $
# Get the vector of 'ind' from 'x.df'
```{r}
#columns
colnames(x.df)
x.df$ind
x.df[,1]
#list
names(newList1)
newList1$a
```

# Modifying Element
```{r}
# Change the element of 'xMat' in the third row and first column to '6'
xMat[3,1] <- 6
# Replace the seconf column of 'x.mat' by 0's
z.mat[,2] <- 0
```
