---
title: "CN334 Data Analytics Lab Assignment"
author: "Aakriti Sharma"
output:
  pdf_document: default
  html_document: default
date: "2026-03-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1. Data Exploration and Descriptive Statistics: #a. Program on data reading and writing using csv file. #b. Load a dataset of your choice and perform initial exploratory data analysis.

```{r}
data <- data.frame(
  Name = c("A", "B", "C", "D", "E"),
  Age = c(23, 25, 22, 24, 26),
  Marks = c(85, 90, 78, 88, 92)
)

write.csv(data, "student_data.csv", row.names = FALSE)

read_data <- read.csv("student_data.csv")

print(read_data)
data(iris)

head(iris)

str(iris)

summary(iris)

sum(is.na(iris))

```
#c. Calculate descriptive statistics (mean, median, standard deviation) for a specific variable in the dataset.

```{r}
mean_val <- mean(iris$Sepal.Length)

median_val <- median(iris$Sepal.Length)

sd_val <- sd(iris$Sepal.Length)


cat("Mean:", mean_val, "\n")
cat("Median:", median_val, "\n")
cat("Standard Deviation:", sd_val, "\n")
```
#d. Visualize the distribution of a numerical variable using a histogram or box plot.
```{r}
boxplot(iris$Sepal.Length,
        main = "Boxplot of Sepal Length",
        col = "lightgreen")
```

## 2. Data Cleaning and Preprocessing:  
#a. Handle missing values in a dataset using appropriate techniques (e.g., imputation, deletion)

```{r}
data(mtcars)
head(mtcars)

# Introduce missing values
mtcars$mpg[1] <- NA
mtcars$hp[5] <- NA
# Count missing values
colSums(is.na(mtcars))
# Remove rows with missing values
mtcars_clean <- na.omit(mtcars)
# Replace NA with mean
mtcars$mpg[is.na(mtcars$mpg)] <- mean(mtcars$mpg, na.rm = TRUE)
mtcars$hp[is.na(mtcars$hp)] <- mean(mtcars$hp, na.rm = TRUE)
```

#b. Identify and handle outliers in a dataset using visualization and statistical methods.

```{r}
# Boxplot to detect outliers
boxplot(mtcars$mpg, main = "Boxplot of MPG")

# Calculate IQR
Q1 <- quantile(mtcars$mpg, 0.25)
Q3 <- quantile(mtcars$mpg, 0.75)
IQR_val <- IQR(mtcars$mpg)

# Define bounds
lower_bound <- Q1 - 1.5 * IQR_val
upper_bound <- Q3 + 1.5 * IQR_val

# Identify outliers
outliers <- mtcars$mpg[mtcars$mpg < lower_bound | mtcars$mpg > upper_bound]
print(outliers)
#Remove Outliers
mtcars_no_outliers <- mtcars[mtcars$mpg >= lower_bound & mtcars$mpg <= upper_bound, ]
#Replace Outliers with Median
median_val <- median(mtcars$mpg)

mtcars$mpg[mtcars$mpg < lower_bound | mtcars$mpg > upper_bound] <- median_val
```

#c. Perform data normalization or standardization on a numerical variable. 

```{r}
#Normalization (Min-Max Scaling)
normalize <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}

mtcars$mpg_norm <- normalize(mtcars$mpg)

head(mtcars$mpg_norm)
```